Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-09-23 thesis/article.tex
Thesis: Refactors split function section into chapter, reaches main theorem
Many fixes and rewrites, small improvements in other sections

% Commands for wrapping properly common expressions.
\newcommand{\indeq}[1]{\stackrel{\text{#1}}{=}}
\newcommand{\RightarrowArg}[1]{\stackrel{#1}{\Longrightarrow}}
\newcommand{\LeftrightarrowArg}[1]{\stackrel{#1}{\Leftrightarrow}}
\newcommand{\NE}{\mathrm{N.E.}}
\newcommand{\as}{\mathrm{\alpha_s}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator*{\argmax}{arg\,max}
% \newcommand{\Exp}{\mathrm{Exp}}
% \newcommand{\Expect}{{\rm I\kern-.3em E}}

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(x(t)) = \ell_i(x_i) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Example \ref{example:a_0_0}, allowing many players close to $0$ breaks the pricing game.

With the above we can again define a Nash equilibrium for the game if $\forall i \in N$ and $\forall p \in [0, 1]$ it holds that
\[c_{f(p)}^p(x(t)) \leq c_i^p(x(t))\]
Existence of an equilibrium is guaranteed by the more general results of Schmeidler \cite[Thm 2]{1973JSP.....7..295S}, while uniqueness is covered by Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220}; for a more detailed coverage of those look at the definitions and propositions of Cole et al. \cite[\S2]{10.1145/780542.780618}.
Therefore can extend the definition for $x(t)$ similarly as in the homogeneous case.

@@

Focusing on the model that will be used in this thesis, simplify our network by only having two links but consider that the flow is acting in a heterogeneous way.
The model can thus be defined as the tuple $(\ell, \alpha)$, where $\ell = (\ell_1, \ell_2)$ the latency functions and $\alpha$ the distribution function.
[TODO: Add figure]

\chapter{2-link parallel networks}

We start with the simple case of 2-link parallel networks.
[TODO: Add figure]
In this simple setup a user of the flow

\section{Time-money sensitivity split}

Consider a 2-link parallel network with latency functions $l_1(x), l_2(x)$ and tolls $t=(t_1, t_2)$ assigned to them.
Without loss of generality we assume $t_1 > t_2$ which in turn means that for the resulting equilibrium flow $x = (x_1(t), x_2(t))$ it holds that $l_1(x_1(t)) < l_2(x_2(t))$.
This equilibrium is guaranteed to exist according to Harks et al. \cite{Harks_2019}.
For the 2 edges of the game we introduce the term time-money sensitivity split $\as(t)$ as the value for which if any player $p$ has $\alpha(p)=\as(t)$ then that player sees the $2$ edge total latencies as equal.
More specifically, $\as(t)$ is the value of the distribution function $\alpha$ for which it holds
\[l_1(x_1(t)) + \as(t) t_1 = l_2(x_2(t)) + \as(t) t_2\]
Solving for $\as$ we get
\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\\
Before we get into a more formal definition and description of $\as$, it'd help to first discuss the nature of the split that $\as$ captures.
Regardless of whether there exists a player $p$ such that $\alpha(p) = \as(t)$, it still holds that
\begin{itemize}
	\item if $\alpha(p) < \as$ then $l_1(x_1(t)) + \alpha(p) t_1 < l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $1$
	\item if $\alpha(p) > \as$ then $l_1(x_1(t)) + \alpha(p) t_1 > l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is on edge $2$
	\item if $\alpha(p) = \as$ then $l_1(x_1(t)) + \alpha(p) t_1 = l_2(x_2(t)) + \alpha(p) t_2$, thus $p$ is either on edge $1$ or $2$
\end{itemize}
Remember that we have assumed that $t_1 > t_2 \Rightarrow l_1(x_1(t)) < l_2(x_2(t))$.
Therefore the flow passing through the faster edge $1$ is contained with players where $\alpha(p) \le \as(t)$, while the flow in the cheaper edge $2$ is contained with players where $\alpha(p) \ge \as(t)$.
Since the players are sorted in increasing order according to $\alpha$, it helps to view the players in the first case as the "lower" part of the split and respectively the players in the second case as the "higher" part of the split.
[TODO: Alternative splits: rushed-relaxed, rich-poor]
Table \autoref{table:split_summary} provides of a summary for each part of the split

\begin{table}[h!]
	\centering
	\caption{Summary of properties for the sensitivity split.}
	\begin{tabular}{| c || c | c |}
		\hline
		& $\alpha(p) \le \as(t)$ & $\alpha(p) \ge \as(t)$ \\ \hline
		sensitivity & time $\ge$ money & time $\le$ money \\ \hline
		latency & lower $(l_2)$ & higher $(l_1)$ \\ \hline
		toll & higher $(t_1)$ & lower $(t_2)$ \\ \hline
		edge & low-latency, high-toll $(1)$ & high-latency, low-toll $(2)$ \\ \hline
		$t_1 - t_2$ & lower & higher \\ \hline
		split & lower & higher \\ \hline
	\end{tabular}
	\label{table:split_summary}
\end{table}

Finally to acknowledge that we have arbitrarily handled any players with $\alpha(p) = \as(t)$.
Those players, if any, even though they see both edges with equal total cost, in the optimal flow $x$ they have settled in either edge $1$ or $2$.
We investigate more in depth the relation of $\as$ to the flow in [TODO: add ref].
\\[12pt]
We can now formally define $\as$ and prove some of its properties.
\begin{definition}
	Given a 2-link parallel heterogeneous game with latency functions $l_1, l_2$, we define the time-money sensitivity split function $\as: \R^2_+ \rightarrow (0, +\infty)$ where for given tolls $t = (t_1, t_2)$
	\[\as(t) = \frac{l_2(x_2(t)) - l_1(x_1(t))}{t_1 - t_2}\]
\end{definition}
The function is well defined since the given tolls $t$ result in a single optimal flow $x(t) = (x_1(t), x_2(t))$ \cite{Harks_2019}.

We start with some basic properties.
\begin{lemma}
	\label{lemma:split_basic}
	Given a 2-link parallel heterogeneous game with distribution function $\alpha$ and time-money sensitivity split function $\as$, the following hold:
	\begin{enumerate}[(i)]
		\item $\as(t) > 0$
		\item if $x_1(t), x_2(t) > 0$ then $\as(t) \in [\alpha(0), \alpha(1)]$
		\item if $0 < l_1(x_1(t)) < l_2(x_2(t))$ then $\alpha(x_1(t)) \le \as(t) \le \alpha(x_1(t) + \epsilon), \forall \epsilon > 0$ arbitrarily small
	\end{enumerate}
\end{lemma}

\begin{proof}
	$ $ % Just to make a line break after "Proof".
	\begin{enumerate}[(i)]
		\item Obviously $\as(t) \ne 0$ since $\as(t) = 0 \Rightarrow l_1(x_1(t)) = l_2(x_2(t))$ which is impossible in a heterogeneous setup.
		Then we only need to notice that $l_1(x_1(t)) < l_2(x_2(t)) \iff t_1 > t_2$ and we get $\as(t) > 0$ from definition.
		\item Since $x_1(t), x_2(t) > 0$ then $\exists p_1, p_2 \in [0, 1]$ each on a different edge, for which we assume w.l.o.g. that $\alpha(p_1) \le \alpha(p_2)$.
		It holds that $\alpha(p_1) \le \as(t) \le \alpha(p_2)$ so with $\alpha$ being non-decreasing it follows
		\[\alpha(0) \le \alpha(p_1) \le \as(t) \le \alpha(p_2) \le \alpha(1) \Rightarrow \as(t) \in [\alpha(0), \alpha(1)]\]
		\item We know that $x_1(t) < x_2(t)$ means that edge 1 has the lower split.
		Therefore we have
		\[\alpha(p) \le \as(t), \forall p \in [0, x_1(t)] \Rightarrow \alpha(x_1(t)) \le \as(t)\]
		Likewise for edge 2 we have
		\[\as(t) \le \alpha(p), \forall p \in (x_1(t), 1] \Rightarrow \as(t) \le \alpha(x_1(t) + \epsilon), \forall \epsilon \in (0, 1 - x_1(t))\]
	\end{enumerate}
\end{proof}

The next lemma will help us describe how $\as(t)$ is changing in relation to toll difference (and in extend to latency difference).
\begin{lemma}
	Given a 2-link parallel heterogeneous game with time-money sensitivity split function $\as$ and tolls $t, t^\prime$ such that $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} > 0$, the following hold:
	\begin{enumerate}[(i)]
		\item $\as(t) \ge \as(t^\prime)$ iff $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \le 1$
		\item $\as(t) \le \as(t^\prime)$ iff $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \ge 1$
	\end{enumerate}
\end{lemma}

\begin{proof}
	Before we get into the propositions, notice that the condition
	\begin{equation}
		\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} > 0
		\label{eq:split_property_cond_tolls_0}
	\end{equation}
	ensures that the two tolls provide splits with same edges in their respective lower and higher sections.
	We show this for $t_1 > t_2$ and likewise it also holds for $t_1 < t_2$.
	\[t_1 > t_2 \Leftrightarrow t_1 - t_2 > 0 \LeftrightarrowArg{(\ref{eq:split_property_cond_tolls_0})} t_1^\prime - t_2^\prime > 0 \Leftrightarrow t_1^\prime > t_2^\prime\]
	%Also note that this property extends to the latencies, giving us
	%\begin{equation}
	%	\frac{l_1(x_1(t)) - l_2(x_2(t))}{l_1(x_1^\prime(t)) - l_2(x_2^\prime(t))} > 0
	%	\label{eq:split_property_cond_latencies_0}
	%\end{equation}
	Now moving on to the prepositions, we focus on proving $(i)$ in the right-to-left direction; then the same steps can be followed backwards to prove left-to-right.
	Finally proving $(ii)$ will be much easier as there are many steps symmetric to the ones for $(i)$.

	\paragraph{$(i)$}
	Begin by noticing that $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \le 1 \Leftrightarrow |t_1 - t_2| \le |t_1^\prime - t_2^\prime|$ (opposite direction is ensured by $(\ref{eq:split_property_cond_tolls_0})$), i.e. the toll difference is increasing.
	Assume w.l.o.g. that $x_1 < x_2$.

	If $x_1(t) = 0$ then
	\[\as(t) = \frac{l_2(1)}{t_2 - t_1}, \as(t^\prime) = \frac{l_2(1)}{t_2^\prime - t_1^\prime}\] 
	so obviously $t_2 - t_1 \le t_2^\prime - t_1^\prime \Rightarrow \as(t) \ge \as(t^\prime)$.

	If $x_1(t) > 0$ then we know [TODO: do we?] that the latency difference might also increase, with flow from the lower latency edge, if any, moving towards the slower one, i.e. $x_1(t) \ge x_1(t^\prime)$.
	%\[|t_1 - t_2| \le |t_1^\prime - t_2^\prime| \Leftrightarrow |l_1(x_1(t)) - l_2(x_2(t))| \le |l_1(x_1(t^\prime)) - l_2(x_2(t^\prime))|\]
	%\begin{equation}
	%	|l_1(x_1(t)) - l_2(x_2(t))| \le |l_1(x_1(t^\prime)) - l_2(x_2(t^\prime))| \LeftrightarrowArg{(\ref{eq:split_property_cond_latencies_0})} \frac{l_1(x_1(t)) - l_2(x_2(t))}{|l_1(x_1(t^\prime)) - l_2(x_2(t^\prime))|} \le 1
	%	\label{eq:split_property_cond_latencies_1}
	%\end{equation}
	If no flow moves then we can easily show that $\as(t) \ge \as(t^\prime)$ similarly to the case with $x_1(t) = 0$.
	Otherwise we have $x_1(t) > x_1(t^\prime) \Rightarrow \exists \epsilon > 0: x_1(t) = x_1(t^\prime) + \epsilon$ and by carefully applying property $\ref{lemma:split_basic}(ii)$ we get
	\[\as(t^\prime) \le \alpha(x_1(t^\prime) + \epsilon) = \alpha(x_1(t)) \le \as(t)\]

	\paragraph{$(ii)$}
	Symmetrically to $(i)$ in this case the toll difference is decreasing, and therefore flow from the higher latency edge will move towards the lower one, i.e. $x_1(t) \le x_1(t^\prime)$ (assuming again $x_1 < x_2$).
	With a similar argument we then get $\as(t) \le \as(t^\prime)$.
\end{proof}

The next lemma gives us some insight of the behavior of $\as(t)$ around its edge values.
\begin{lemma}
	\label{lemma:split_bounds}
	Given a 2-link parallel heterogeneous game with distribution function $\alpha$ and time-money sensitivity split function $\as$, the following hold:
	\begin{enumerate}[(i)]
		\item if $\alpha(0) > 0$ then $\as(t) < \alpha(0) \Rightarrow x(t) = (0, 1)$ or $(1, 0)$
		\item if $x_1(0), x_2(0) > 0$ and $\alpha(0) = 0$ with $0$ an accumulation point, then $x_1(t), x_2(t) > 0, \forall t \in \R_+^2$
		\item $\lim_{|t_1 - t_2| \rightarrow +\infty}\as(t) = 0$
		\item $\lim_{t_1 - t_2 \rightarrow 0^-} \as(t) = \alpha(x_1), \lim_{t_1 - t_2 \rightarrow 0^+} \as(t) = \alpha(x_2)$, where ($x_1, x_2)$ the optimal flow under no tolls
	\end{enumerate}
\end{lemma}

\begin{proof}
	$ $
	\begin{enumerate}[(i)]
		\item Direct consequence from the fact that all players $p$ with $\as(t) < \alpha(p)$ are in the low latency edge.
		\item Since $\alpha(0) = 0$ and $0$ an accumulation point, then $\forall \delta > 0 \exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \delta$.
		Therefore $\forall t \in \R_+^2$ if $\delta = \as(t)$ then $\exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \as(t)$, which means that the lower latency edge will always have flow some $\epsilon > 0$, and therefore $x_1(t), x_2(t) > 0$.
		\item We only need to notice that $|l_1(x_1(t)) - l_2(x_2(t))| \le \max\{l_1(1), l_2(1)\} = L$ and we find
		\[\lim_{|t_1 - t_2| \rightarrow +\infty}\as(t) = \lim_{|t_1 - t_2| \rightarrow +\infty}\frac{|l_1(x_1(t)) - l_2(x_2(t))|}{|t_1 - t_2|} \le \lim_{|t_1 - t_2| \rightarrow +\infty}\frac{L}{|t_1 - t_2|} = 0\]
		and since $\as(t) > 0$ we get $\lim_{|t_1 - t_2| \rightarrow +\infty}\as(t) = 0$.
		\item As $|t_1 - t_2| \rightarrow 0$ then also $|l_1(x_1(t)) - l_2(x_2(t))| \rightarrow 0$ as the game approaches the homogeneous case.
		Then flow on the low latency edge will define the limit of $\as(t)$.
			\begin{itemize}
				\item For $t_1 > t_2$ we get lower latency edge $1$ with flow $x_1(t) \rightarrow x_1$ and therefore
				\[\lim_{t_1 - t_2 \rightarrow 0^-} \as(t) = \alpha(x_1)\]
				\item For $t_1 < t_2$ we get lower latency edge $2$ with flow $x_2(t) \rightarrow x_2$ and therefore
				\[\lim_{t_1 - t_2 \rightarrow 0^+} \as(t) = \alpha(x_2)\]
			\end{itemize}
	\end{enumerate}
\end{proof}

From the above lemma we can also observe the following

\begin{corollary}
	Given a 2-link parallel heterogeneous game with distribution function $\alpha$ and time-money sensitivity split function $\as$ and $x = (x_1, x_2)$ the optimal flow in the game without tolls, then $\as(t) \le \max\{\alpha(x_1), \alpha(x_2)\}$.
\end{corollary}

>>>>

% Commands for wrapping properly common expressions.
\newcommand{\indeq}[1]{\stackrel{\text{#1}}{=}}
\newcommand{\RightarrowArg}[1]{\stackrel{#1}{\Rightarrow}}
\newcommand{\LeftrightarrowArg}[1]{\stackrel{#1}{\Leftrightarrow}}
\newcommand{\NE}{\mathrm{N.E.}}
\newcommand{\as}{\mathrm{\alpha_s}}
\newcommand{\R}{\mathbb{R}}
\DeclareMathOperator*{\argmax}{arg\,max}
% \newcommand{\Exp}{\mathrm{Exp}}
% \newcommand{\Expect}{{\rm I\kern-.3em E}}

@@

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(t) = \ell_i(x_i(t)) + \alpha(p) \cdot t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Finally, we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Example \ref{example:a_0_0}, allowing many players close to $0$ breaks the pricing game.

We can therefore define an instance of a Heterogeneous Pricing Game as the tuple $(N, \ell, \alpha)$ with $N = \{1, 2, \dots, n\}$, $\ell = (\ell_i)_{i \in N}$ and $\alpha$ the distribution function.
Then with all the above we can again define a Nash equilibrium for the game if $\forall i \in N$ and $\forall p \in [0, 1]$ it holds that
\[c_{f(p)}^p(t) \leq c_i^p(t)\]
Existence of an equilibrium is guaranteed by the more general results of Schmeidler \cite[Thm 2]{1973JSP.....7..295S}, while uniqueness is covered by Milchtaich \cite[Prop 3.3]{doi:10.1287/moor.25.3.349.12220}; for a more detailed coverage of those look at the definitions and propositions of Cole et al. \cite[\S2]{10.1145/780542.780618}.
Therefore can extend the definition for $x(t)$ similarly as in the homogeneous case.

Finally, we will prove here a general property of parallel networks which capture the relationship between latencies and tolls.
It is trivial for homogeneous games that higher tolls create lower latencies on the respective edges, but for the heterogeneous case it is not straightforward so we'll prove it.

\begin{lemma}
	\label{lemma:latencies_tolls}
	For a Heterogeneous Parallel Game $(N, \ell, \alpha)$ with tolls $t$ and Nash equilibrium for $t$ flow $x(t)$, it holds that for all $i, j \in N$ with $x_i(t), x_j(t) > 0$
	\begin{enumerate}[(i)]
		\item $\ell_i(x_i(t)) < \ell_j(x_j(t))$ iff $t_i > t_j$
		\item $\ell_i(x_i(t)) = \ell_j(x_j(t))$ iff $t_i = t_j$
		\item $\ell_i(x_i(t)) > \ell_j(x_j(t))$ iff $t_i < t_j$
	\end{enumerate}
\end{lemma}

\begin{proof}
	We will prove $(i)$ and the rest follow similarly.
	For the left-to-right direction we assume that $\ell_i(x_i(t)) < \ell_j(x_j(t))$.
	If now $t_i \le t_j$ then for all players $p$ we get $\ell_i(x_i(t)) + \alpha(p) \cdot t_i < \ell_j(x_j(t)) + \alpha(p) \cdot t_j \Rightarrow c_i^p(t) < c_j^p(t)$.
	Therefore there exist players $p_j$ on link $j$ for which $c_{f(p_j)}^{p_j}(t) = c_j^{p_j}(t) > c_i^{p_j}(t)$, which is a contradiction since $x(t)$ is a Nash equilibrium.
	The opposite direction is also proved similarly.
\end{proof}

@@

Focusing on the model that will be used in this thesis, we simplify our network by only having $2$ links, and then consider that the flow is acting in a heterogeneous way.
The model can thus be defined as the tuple $([2], \ell, \alpha)$, where $[2] = \{1, 2\}$ the links, $\ell = (\ell_1, \ell_2)$ the latency functions and $\alpha$ the distribution function.
[TODO: Add figure]
When discussing properties of the model where the pricing game is not relevant, we will refer to a $([2], \ell, \alpha)$ instance as a toll game, and otherwise as a pricing game.

\chapter{Money sensitivity split}

\section{Introduction}

Consider a $([2], \ell, \alpha)$ toll game such that $x_1(0), x_2(0) > 0$ with tolls $t = (t_1, t_2)$ assigned to them and  $x(t) = (x_1(t), x_2(t))$ the Nash equilibrium for $t$.
Without loss of generality we assume $t_1 > t_2$ which from Lemma \ref{lemma:latencies_tolls} follows that $\ell_1(x_1(t)) < \ell_2(x_2(t))$.
For the $2$ edges of the game we introduce the term money sensitivity split $\as(t)$ as the value of $\alpha$ for which if any player $p$ has $\alpha(p)=\as(t)$ then that player sees the $2$ edges' total latencies as equal.
More specifically, $\as(t)$ is the value of the distribution function $\alpha$ for which it holds
\[\ell_1(x_1(t)) + \as(t) \cdot t_1 = \ell_2(x_2(t)) + \as(t) \cdot t_2\]
Solving for $\as$ we get
\[\as(t) = \frac{\ell_2(x_2(t)) - \ell_1(x_1(t))}{t_1 - t_2}\]
\\
Before we get into a more formal definition and description of $\as$, it'd help to first discuss the nature of the split that $\as$ captures.
Regardless of whether there exists a player $p$ such that $\alpha(p) = \as(t)$, it still holds that
\begin{itemize}
	\item if $\alpha(p) < \as(t)$ then $\ell_1(x_1(t)) + \alpha(p) \cdot t_1 < \ell_2(x_2(t)) + \alpha(p \cdot) t_2$, thus $p$ is on edge $1$
	\item if $\alpha(p) > \as(t)$ then $\ell_1(x_1(t)) + \alpha(p) \cdot t_1 > \ell_2(x_2(t)) + \alpha(p) \cdot t_2$, thus $p$ is on edge $2$
	\item if $\alpha(p) = \as(t)$ then $\ell_1(x_1(t)) + \alpha(p) \cdot t_1 = \ell_2(x_2(t)) + \alpha(p) \cdot t_2$, thus $p$ is either on edge $1$ or $2$
\end{itemize}
Remember that we have assumed that $t_1 > t_2 \Rightarrow \ell_1(x_1(t)) < \ell_2(x_2(t))$.
Therefore the flow passing through the faster edge $1$ is contained with players where $\alpha(p) \le \as(t)$, while the flow in the cheaper edge $2$ is contained with players where $\alpha(p) \ge \as(t)$.
Since the players are sorted in increasing order according to $\alpha$, it helps to view the players in the first case as the "lower" part of the split and respectively the players in the second case as the "upper" part of the split.
If it helps with the intuition, alternative split labels could be rushed-relaxed or rich-poor.
Table \autoref{table:split_summary} provides of a summary for each part of the split

\begin{table}[h!]
	\centering
	\caption{Summary of properties for the sensitivity split.}
	\begin{tabular}{| c || c | c |}
		\hline
		& $\alpha(p) \le \as(t)$ & $\alpha(p) \ge \as(t)$ \\ \hline
		sensitivity & time $\ge$ money & time $\le$ money \\ \hline
		latency & lower $(\ell_2)$ & higher $(\ell_1)$ \\ \hline
		toll & higher $(t_1)$ & lower $(t_2)$ \\ \hline
		link/edge & low-latency, high-toll $(1)$ & high-latency, low-toll $(2)$ \\ \hline
		$t_1 - t_2$ & lower & higher \\ \hline
		split & lower & upper \\ \hline
	\end{tabular}
	\label{table:split_summary}
\end{table}

Finally to acknowledge that we have arbitrarily handled any players with $\alpha(p) = \as(t)$.
Those players, if any, even though they see both edges with equal total cost, in the optimal flow $x$ they have settled in either edge $1$ or $2$.
We investigate more in depth the relation of $\as$ with the flow in the following Lemmas.
\\[12pt]
We can now formally define $\as$ and prove some of its properties.

\section{Definitions}

\begin{definition}
	\label{definition:split_function}
	Given a $([2], \ell, \alpha)$ toll game with $x_1(0), x_2(0) > 0$, tolls $t = (t_1, t_2), t_1 \ne t_2$ and $x(t) = (x_1(t), x_2(t))$ the Nash equilibrium for $t$, we define the \textit{money sensitivity split function} $\as: \R_+^2 \setminus \{(a, a) | a \in \R_+\} \rightarrow (0, +\infty)$ as follows 
	\[\as(t) = \frac{\ell_2(x_2(t)) - \ell_1(x_1(t))}{t_1 - t_2}\]
\end{definition}

We will also give a more formal definition to the lower-upper split intuition we described above.
\begin{definition}
	\label{definition:split_lower_upper}
	Given a $([2], \ell, \alpha)$ toll game with $x_1(0), x_2(0) > 0$, tolls $t \ne 0$ and $x(t)$ the Nash equilibrium for $t$, and also assuming w.l.o.g. that $t_1 > t_2$, we define as \textit{lower split} the (possibly zero) flow $x_1(t)$ and as \textit{upper split} the remaining flow $x_2(t)$. 
\end{definition}

\begin{corollary}
	\label{corollary:split_to_alpha}
	Given a $([2], \ell, \alpha)$ toll game with $x_1(0), x_2(0) > 0$ and tolls $t \ne 0$, it holds that
	\begin{itemize}
		\item $\alpha(p) \le \as(t)$, for all players $p$ in the lower split
		\item $\alpha(p) \ge \as(t)$, for all players $p$ in the upper split
	\end{itemize}
\end{corollary}

\begin{proof}
	Let again w.l.o.g. assume $t_1 > t_2 \LeftrightarrowArg{\ref{lemma:latencies_tolls}} \ell_1(x_1(t)) < \ell_2(x_2(t))$, making by definition \ref{definition:split_lower_upper} a player $p$ in the lower split be in link $1$.
	Since $x$ is a Nash equilibrium we have
	\begin{align*}
		c_1^p(t) &\le c_2^p(t) \Rightarrow \ell_1(x_1(t)) + \alpha(p) \cdot t_1 \le \ell_2(x_2(t)) + \alpha(p) \cdot t_2 \\
		&\RightarrowArg{t_1 > t_2} \alpha(p) \le \frac{\ell_2(x_2(t)) - \ell_1(x_1(t))}{t_1 - t_2} \RightarrowArg{\ref{definition:split_function}} \alpha(p) \le \as(t)
	\end{align*}
	Similarly we can show that $\alpha(p) \ge \as(t)$, for all players $p$ in the upper split.
\end{proof}

At this point in our analysis, we will additionally need the notion of the lowest and highest $\alpha$ value in a flow of players.

In general, since we're working with a Lebesque-measurable flow, it is natural to not bother with distinguishing between a flow $[0, x_1]$ or $[0, x_1)$, as the player $x_1$ can be arbitrarily placed at any link without affecting the game.
In a heterogeneous setting, however, the existence or not of that player matters, as its $\alpha(x_1)$ value might be different from the rest of the players in the flow, depending on how $\alpha$ is defined.
We bypass this issue by instead using set border bounds.

\begin{definition}
	\label{definition:alpha_flow_sets}
	Let $([2], \ell, \alpha)$ a toll game with $x_1(0), x_2(0) > 0$, tolls $t \ne 0$ and $x(t)$ the Nash equilibrium for $t$.
	We define as $A_i(t)$ the set of $\alpha(p)$ values for players $p$ on edge $i$ in the Nash equilibrium for $t$.
	More formally, we define a set function $A_i: \R_+^2 \setminus \{(a, a) | a \in \R_+\} \rightarrow \mathcal{P}(\R_+)$ such that
	\[A_i(t) = \{\alpha(p)|p \text{ in edge } i \textit{ for } x(t) \}\]
	For tolls where $t_1 = t_2$ we cannot define lower and upper split, but we will extend the definition of $A_i$ for those tolls with a sign to specify which split we assume the link to have.
	Therefore we define $A_i^-(0)$ and $A_i^+(0)$ respectively as
	\[A_i^-(0) = \{\alpha(p)|p \in [0, x_i(0))\} \qquad A_i^+(0) = \{\alpha(p)|p \in (x_i(0), 1]\}\]
	%\[A_i^-(0) = \lim_{t_1 - t_2 \rightarrow 0^-}A_i(t) \qquad A_i^+(0) = \lim_{t_1 - t_2 \rightarrow 0^+}A_i(t)\]
\end{definition}

We ignore all $\alpha(x_i(0))$ values in order to avoid any confusion, since they are arbitrarily placed in any edge and their $\alpha$ value could change the sets' properties.
It doesn't affect our analysis, however, since having a split on that player is only possible for $t = 0$ and $\as(0)$ is undefined, plus as we'll show in the next section neither it's limit exists, making the $\alpha$ value of player $x_i(0)$ (in each assumingly split case) irrelevant.

We continue with describing some basic properties around the values of $\as$.

\section{Properties and analysis}

\begin{lemma}
	\label{lemma:split_basic}
	Given a $([2], \ell, \alpha)$ toll game with $x_1(0), x_2(0) > 0$ and tolls $t \ne 0$, the following hold:
	\begin{enumerate}[(i)]
		\item $\as(t) > 0$
		\item if $x_1(t), x_2(t) > 0$ then $\as(t) \in [\alpha(0), \alpha(1)]$
		\item if $\as(t) < \alpha(0)$ or $\as(t) > \alpha(1)$ then $x(t) = (0, 1)$ or $(1, 0)$
		\item if $x_1(t), x_2(t) > 0$ with $\ell_1(x_1(t)) < \ell_2(x_2(t))$ then\\
		$\sup A_1(t) \le \as(t) \le \inf A_2(t)$
	\end{enumerate}
\end{lemma}

\begin{proof}
	$ $ % Just to make a line break after "Proof".
	\begin{enumerate}[(i)]
		\item Trivial application of Lemma \ref{lemma:latencies_tolls}.
		Since we have $t_1 \ne t_2$ it follows that $\ell_1(x_1(t)) \ne \ell_2(x_2(t))$ and therefore $\as(t) \ne 0$.
		Then we only need to notice that $\ell_1(x_1(t)) < \ell_2(x_2(t)) \Leftrightarrow t_1 > t_2$ and we get $\as(t) > 0$ from Definition \ref{definition:split_function}.
		\item Since $x_1(t), x_2(t) > 0$ then $\exists p_1, p_2 \in [0, 1]$ each on a different link, for which we assume w.l.o.g. that $\alpha(p_1) \le \alpha(p_2)$.
		It holds from \ref{corollary:split_to_alpha} that $\alpha(p_1) \le \as(t) \le \alpha(p_2)$ so with $\alpha$ being non-decreasing it follows
		\[\alpha(0) \le \alpha(p_1) \le \as(t) \le \alpha(p_2) \le \alpha(1) \Rightarrow \as(t) \in [\alpha(0), \alpha(1)]\]
		\item Inverse logic statement of $(ii)$ ($(P \rightarrow Q) \rightarrow (\neg Q \rightarrow \neg P)$).
		\item Natural consequence from Corollary \ref{corollary:split_to_alpha} and Definition \ref{definition:alpha_flow_sets}.
	\end{enumerate}
\end{proof}

At this point we should notice that $\as$, similarly to $x$, depends only on the toll difference $t_1 - t_2$ and whether $t_1 < t_2$ or otherwise.
In order to emphasise on this, let's point out that we could alternatively have defined both $x$ and $\as$ as single variable functions with a domain of values $t_d = t_1 - t_2 \in \R$ and function values as such
\[
	x(t_d) =
	\begin{cases*}
		x((t_d, 0)) & $t_d \ge 0$ \\
		x((0, -t_d)) & $t_d < 0$ \\
	\end{cases*}
\]
We will not utilise this definition as is in this work, however displaying it helps with the intuition that the flow, and consequently $\as$, actually changes as the toll difference does in $\R$.

With that in mind, we will investigate the monotonicity of $\as(t)$ in relation to toll differences (and in extension to latency differences).
For this we will need to compare different tolls that create lower-upper splits in the same edges, or in other words for tolls $t, t^\prime$ to hold that if $t_1 > t_2$ then also $t_1^\prime > t_2^\prime$ and vice-versa.
We will capture this with the following condition
\begin{equation}
	\label{eq:split_property_cond_tolls_0}
	\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} > 0
\end{equation}
where indeed if $t_1 > t_2$ then
\[t_1 > t_2 \Leftrightarrow t_1 - t_2 > 0 \LeftrightarrowArg{(\ref{eq:split_property_cond_tolls_0})} t_1^\prime - t_2^\prime > 0 \Leftrightarrow t_1^\prime > t_2^\prime\]
and likewise we also get $t_1 < t_2 \Leftrightarrow t_1^\prime < t_2^\prime$.
Also note that this property also holds for the respective latencies.

\begin{lemma}
	\label{lemma:split_monotonicity}
	Given a $([2], \ell, \alpha)$ toll game with $x_1(0), x_2(0) > 0$ and tolls $t, t^\prime \ne 0$ such that $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} > 0$, the following hold:
	\begin{enumerate}[(i)]
		\item if $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \le 1$ then $\as(t) \ge \as(t^\prime)$
		\item if $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \ge 1$ then $\as(t) \le \as(t^\prime)$
		\item if $\as(t) > \as(t^\prime)$ then $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} < 1$
		\item if $\as(t) < \as(t^\prime)$ then $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} > 1$
	\end{enumerate}
\end{lemma}

\begin{proof}
	$ $
	\paragraph{$(i)$}
	Begin by noticing that $\frac{t_1 - t_2}{t_1^\prime - t_2^\prime} \le 1 \Rightarrow |t_1 - t_2| \le |t_1^\prime - t_2^\prime|$, i.e. from $t$ to $t^\prime$ the toll difference is not decreasing, which means that the players in the upper link (i.e. the players who already prefer the higher latency and lower toll link) will still view their link as the best option.
	Therefore only flow from the lower latency edge, if any, might move towards the slower one.
	Assume w.l.o.g. that $t_1 > t_2$, giving us $x_1(t) \ge x_1(t^\prime)$ and $x_2(t) \le x_2(t^\prime)$.
